---
title: "COVID-19 AWARENESS"
author: "Burak Can Onarım"
date: "30 05 2020"
output: html_document
---

# Introduction
**COVID-19**(Coronavirus disease 2019) is an infectious disease caused by **SARS-CoV-2**(severe acute respiratory syndrome coronavirus 2). It was first identified in ***December 2019*** in Wuhan, China, and has since spread globally and also the pandemic era is still ongoing for a long time. So, are people aware of Covid-19 disease? I have created a basic quiz to reach an answer of this question.

***__Note:__*** Language of the quiz is Turkish because all participants are Turks and most of them do not know English.

## Quiz
You can click [here](https://forms.gle/6ikkicK3eNwksTQt5) to reach my quiz. Besides the form link, you can also see a brief summary of quiz in below:

### Demographic Questions
I have asked some questions about demographic information and this list what is included questions:

* *Age*
* *Gender*
* *Marital Status*
* *Education*
* *Monthly Household Income*

### Quiz Questions
The value of correct answers is 10 points. For example, a participant can get 70 points if he/she answers correctly for 7 questions out of 10 questions. Also, you can easily see topics of questions:

1. Lifetime of coronavirus
2. Social distance
3. Vaccine for Covid-19
4. Most risky group
5. Pets can spread coronavirus?
6. Correctness about coronavirus spreading or not spreading in warm climates
7. City where the disease first appeared
8. Symptoms of Covid-19
9. Correctness about effect of using vinegar for coronavirus
10. Example of social distancing

### How I collected data?
First of all, I have created a form using Google Forms. Then, I made a request to my circle of friends and also theirs for answering quiz questions. There were already more over 100 participants in a few hours.

---

# Data Preprocessing
First of processes is downloading dataset as Excel format. The next step is importing dataset from Excel. Also, we can see the column names.
```{r}
library(readxl)
dataset <- read_excel("C:/Users/burak/Desktop/dataset.xlsx")
colnames(dataset)
```

And then, column names should to change with English terms and the questions are not necessary for me. So, I will rename them with question numbers.
```{r}
attach(dataset)
names(dataset)[names(dataset) == "Zaman damgası"] <- "timestamp"
names(dataset)[names(dataset) == "Puan"] <- "total.score"
names(dataset)[names(dataset) == "Yaşınız"] <- "age.range"
names(dataset)[names(dataset) == "Cinsiyetiniz"] <- "gender"
names(dataset)[names(dataset) == "Medeni Durumunuz"] <- "marital.status"
names(dataset)[names(dataset) == "Eğitim Durumunuz"] <- "education"
names(dataset)[names(dataset) == "Aylık Hane Geliri"] <- "house.income"
names(dataset)[names(dataset) == "Yeni koronavirüs vücut dışında ne kadar süre hayatta kalır?"] <- "question.1"
names(dataset)[names(dataset) == "Sosyal mesafe en az ne kadar olmalıdır?"] <- "question.2"
names(dataset)[names(dataset) == "Covid-19 için bir aşı en erken ne zaman piyasaya sunulabilir?"] <- "question.3"
names(dataset)[names(dataset) == "Aşağıdakilerden hangisi Covid-19 hastalığına yakalanırsa kurtulma ihtimali en düşüktür?"] <- "question.4"
names(dataset)[names(dataset) == "Evcil hayvanım bana Covid-19 bulaştırabilir mi?"] <- "question.5"
names(dataset)[names(dataset) == "ABD Başkanı Donald Trump, koronavirüsün yayılmasının daha sıcak havalarda yavaşlayabileceğini söyledi. Covid-19 virüsünün sıcak ve nemli iklimlerde yayılma hızı gerçekten azalır mı?"] <- "question.6"
names(dataset)[names(dataset) == "Covid-19 hastalığının kaynağı olan dünya şehri hangisidir?"] <- "question.7"
names(dataset)[names(dataset) == "Aşağıdakilerden hangisi Covid-19 hastalığının belirtilerinden biri değildir?"] <- "question.8"
names(dataset)[names(dataset) == "Covid-19 hastalığına neden olan koronavirüsün yayılmasına karşı yapılan temizliklerde sirke kullanımı etkili midir?"] <- "question.9"
names(dataset)[names(dataset) == "Aşağıdakilerden hangisi sosyal mesafe kuralına bir örnektir?"] <- "question.10"
```

I think I did everything which can be done for column names. Now, let see new column names.
```{r}
colnames(dataset)
```

Now, let us see data types of all columns.
```{r}
sapply(dataset, typeof)
```

Data type of all columns is ***character*** except __timestamp__ and __total.score__ columns. So these has to convert to factor type for effective data visualization.
```{r}
dataset$age.range <- as.factor(dataset$age.range)

dataset$gender <- as.factor(dataset$gender)

dataset$marital.status <- as.factor(dataset$marital.status)

dataset$education <- as.factor(dataset$education)

dataset$house.income <- as.factor(dataset$house.income)

dataset$question.1 <- as.factor(dataset$question.1)

dataset$question.2 <- as.factor(dataset$question.2)

dataset$question.3 <- as.factor(dataset$question.3)

dataset$question.4 <- as.factor(dataset$question.4)

dataset$question.5 <- as.factor(dataset$question.5)

dataset$question.6 <- as.factor(dataset$question.6)

dataset$question.7 <- as.factor(dataset$question.7)

dataset$question.8 <- as.factor(dataset$question.8)

dataset$question.9 <- as.factor(dataset$question.9)

dataset$question.10 <- as.factor(dataset$question.10)
```

---

After the processes for columns, I want to replace Turkish terms with English ones for demographic informations. Als, I have to arrange factor levels judiciously.
```{r message=FALSE}
library(dplyr)
dataset <- dataset %>% mutate(age.range = recode(age.range, "65 üstü" = "65+"))
dataset$age.range <- factor(dataset$age.range, levels = c("0 - 18", "18 - 25", "25 - 40", "40 - 65", "65+"))

dataset <- dataset %>% mutate(gender = recode(gender, "Erkek" = "Male", "Kadın" = "Female", "Belirtmek İstemiyorum" = "I do not want to state"))
dataset$gender <- factor(dataset$gender, levels = c("Male", "Female", "I do not want to state"))

dataset <- dataset %>% mutate(marital.status = recode(marital.status, "Bekar" = "Single", "Evli" = "Married", "Boşanmış/Dul" = "Divorced/Widowed"))
dataset$marital.status <- factor(dataset$marital.status, levels = c("Single", "Married", "Divorced/Widowed"))

dataset <- dataset %>% mutate(education = recode(education, "İlkokul" = "Elementary School", "Ortaokul" = "Secondary School", "Lise" = "High School", "Önlisans/Lisans" = "Associate/Bachelor's Degree", "Yüksek Lisans veya Üstü" = "Master Degree or above"))
dataset$education <- factor(dataset$education, levels = c("Elementary School", "Secondary School", "High School", "Associate/Bachelor's Degree", "Master Degree or above"))

dataset <- dataset %>% mutate(house.income = recode(house.income, "10.000 TL veya üstü" = "10.000 TL or above"))
dataset$house.income <- factor(dataset$house.income, levels = c("0 - 3.000 TL", "3.000 - 6.000 TL", "6.000 - 10.000 TL", "10.000 TL or above"))
```

Actually I do not need answers. Therefore, I will use **correctness** and **wrongness** of questions instead of answers of them. So now, the process of changing answers is turn to come.
```{r}
dataset <- dataset %>% mutate(question.1 = recode(question.1, "En fazla birkaç gün" = "true", "Bir haftadan fazla" = "false", "Yaklaşık olarak bir ay" = "false", "Bir sene" = "false"))

dataset <- dataset %>% mutate(question.2 = recode(question.2, "Yarım metre" = "false", "Bir metre" = "true", "Beş metre" = "false", "On metre" = "false"))

dataset <- dataset %>% mutate(question.3 = recode(question.3, "İki ay sonra" = "false", "Dört ay sonra" = "false", "Altı ay sonra" = "false", "Bir sene sonra" = "true"))

dataset <- dataset %>% mutate(question.4 = recode(question.4, "Çocuklar" = "false", "Hamileler" = "false", "65 yaş ve üstü kişiler" = "true", "Gençler" = "false"))

dataset <- dataset %>% mutate(question.5 = recode(question.5, "Evet" = "false", "Hayır" = "true"))

dataset <- dataset %>% mutate(question.6 = recode(question.6, "Evet" = "false", "Hayır" = "true"))

dataset <- dataset %>% mutate(question.7 = recode(question.7, "Wuhan" = "true", "Hong Kong" = "false", "Milano" = "false", "New York" = "false"))

dataset <- dataset %>% mutate(question.8 = recode(question.8, "Solunum sıkıntısı" = "true", "Ateş" = "false", "Hapşırık" = "true", "Kuru öksürük" = "false"))

dataset <- dataset %>% mutate(question.9 = recode(question.9, "Evet" = "false", "Hayır" = "true"))

dataset <- dataset %>% mutate(question.10 = recode(question.10, "Kalabalık yerlere gitmeyi bırakmak" = "true", "Aynı evde yaşadığın kişilerle konuşmayı bırakmak" = "false", "Telefonla konuşmayı bırakmak" = "false", "Televizyon izlemeyi bırakmak" = "false"))
```

---

Finally, let see the new dataframe for first six rows:

```{r}
library(knitr)
kable(head(dataset), format = "markdown", align = "c")
```

---

## Data Visualization
```{r, echo=FALSE}
detach(dataset)
attach(dataset)
```

I want to see some identifier information about score of participants. So, I am goint to use **_boxplot_** and **_histogram_**.

```{r}
boxplot(total.score)
```

According to this boxplot, most of participants get high scores. First qurtile is ***70*** and third quartile is ***90***. Also, the average of scores is seem to be ***80***. Let me check the correct average:
```{r}
mean(total.score)
```

As you can see that is in the preceding output there is an average value for total.score and it is an integer which is valued of 80.

### Some Interpretation about *total.score*
* The minimum score is *40* and maximum is *100*. So, the range is actually wide. 
  + This result shows that there are participants who are various.
  + It shows that also there are not participants who are very unconscious.

* In the boxplot, the distance between maximum point and box is shorter than distance between minimum point and box.
  + This result shows that many participants get a high scores.

* My final comment is about number of participants. I've seen that most of participants get high scores when I look the boxplot.

Let us see more clearly using histogram for *total.score*.

### *Data Visualization(cont.)*

Now, I will utilize function of hist().
```{r}
hist(total.score, main = "Scores of Participants",
     xlab = "Score", ylab = "Participants", col = "steelblue",
     border = "steelblue", breaks = 100)
```

As it was expected, there are no more participants who get low grades.

Also,

* The number of participants who gets average points are more over comparing to others.

* The number of participants who gets 70 or more points(i.e. high points) is much.

### *Based on Demographics*

Now, I want to see that demographic informations for participants:
```{r}
library(ggplot2)
library(ggthemes)
ggplot(dataset) + geom_count(aes(x = house.income, y = age.range), color = "steelblue") +
  facet_grid(marital.status ~ gender) +
  labs(tag = "First Plot",
       title = "Demographic Informations",
       subtitle =  "but not included info of education",
       x = "Household Income", y = "Age",
       caption = "Created by Burak Can Onarım in 31.05.2020") +
  theme(axis.ticks = element_blank(),
        axis.text.x = element_text(angle=90, hjust=1))
```

There are many inference:

* Most of participants have stated their gender information.
  + All participants who have **not** stated their gender are *married*, age is *between 25 and 40*, also house income is *more than 10.000 TL or equal*.
  + Number of female participants is seem to be greater than males.
  + The house income of all male participants is less than 6.000 TL or equal. On the other hand, house income of females is various.
  
* Most of participants are **single**.
  + There are no married males who have house income more than *6.000 TL*.
  + As it was expected, age of married participants is minimum *18*; but not more than 65.
  
* Most of participants are **40 years old or less than 40**.
  + All of 65 or more years old participants are *female*, *single* and their house income is *highest option*.
  + Single and female participants who are between 25 and 65 years old did not participated.

* The number of participants decreases as the house income increases.
  + Most of participants have *6.000 TL or less than 6.000 TL* house income.
  + All of male and married participants have not selected highest option for house income.
  
In addition, many inferences can be made using this plot.

---

### Score ~ Income

Would the total.score be increased when household income increase? Let us see using the boxplot:
```{r, warning=FALSE,message=FALSE}
ggplot(dataset) + geom_boxplot(mapping = aes(x = house.income, y = total.score)) +
  scale_y_continuous(breaks = seq(40, 100, 10)) +
  labs(tag = "Second Plot",
       title = "Scores based on Household Income",
       subtitle =  "using boxplot",
       x = "Household Income", y = "Score",
       caption = "Created by Burak Can Onarım in 31.05.2020")
```

As I expected, awareness of participants increases when their incomes increases. The awareness about Covid-19 is same for the participants who earn *0 - 3.000 TL* and *3.000 - 6.000 TL* at monthly.

---

### Score ~ Age range and Gender

I want to show that score distribution based on age range and gender together:
```{r}
ggplot(dataset) + geom_bar(mapping = aes(x = total.score), width = 5) +
  facet_wrap(age.range ~ gender, nrow = 2) +
  labs(tag = "Third Plot",
       title = "Scores based on Age and Gender",
       subtitle =  "using barplot",
       x = "Score", y = "Count",
       caption = "Created by Burak Can Onarım in 31.05.2020")
```

---

### Scores ~ Education

Now, I want to show that scores with educations. Does the score increase as education level ups?
```{r}
ggplot(dataset, aes(total.score, fill = education)) + geom_bar(width = 5) +
  scale_y_continuous(breaks = seq(0, 60, 10)) +
  labs(tag = "Fourth Plot",
       title = "Scores based on Education",
       subtitle =  "using barplot",
       x = "Score", y = "Count",
       caption = "Created by Burak Can Onarım in 31.05.2020")
```

* For master degree or above;
  + Their minimum score is *60* and maximum is *70*. So, they have not much awareness score about covid-19.
  + I think they should learn more stuffs for protection from coronavirus.
  
* For associate or bachelor's degree;
  + They constitute the vast majority of the participants. So, their awareness is like all participants.
  + For most of them, their scores are *80* or above.
  
* For high school;
  + Some few have *100* point but the rest have chosen at least one wrong answer.
  + They should learn more stuffs.
  
* For secondary school;
  + There are not many people who get *90* or more points.
  + They should learn more stuff, too.
  
* For elementary school;
  + They do not get less than *60* points.
  + They are equal-distributed for any score higher than *60*.
  
  
**MY RESULT --> ** Education level is not effective for awareness about Covid-19 disease. So, schools are not teaching anything about it. For awareness, people have to learn by themselves.

---

### Scores ~ Marital Status

Final plot is about effects of marital status on awareness score.

```{r, warning=FALSE}
ggplot(dataset, aes(total.score)) + geom_bar(width = 5) +
  facet_wrap(~ marital.status) +
  scale_y_continuous(breaks = seq(0, 60, 10)) +
  labs(tag = "Fifth Plot",
       title = "Scores based on Marital Status",
       subtitle =  "using barplot",
       x = "Score", y = "Count",
       caption = "Created by Burak Can Onarım in 31.05.2020")
```

* For single participants;
  + Many of them get *80* or more points. So, their awareness is not bad.
  + Nevertheless, a couple of single participants could not get even *60* point. They have to improve themselves about Covid-19 disease.

* For married participants;
  + Like singles, many of them get *80* or more points. So, their awareness is good.
  + They are successful, although there is not much participation.

* For divorced or widowed participants;
  + There are only a few divorced or widowed participants.
  + All of them got *70* points.

---

## Confidence Intervals

I will compare confidence intervals of mean total.score for high educated(i.e. education level is higher than high school) AND low educated(i.e. education level is lower than high school or equal)

### Pre-processing

I want to create two dataframes based on education levels. So,

```{r}
# All participants whose education level is higher than high school;
high.educated <- dataset[!(dataset$education == "Elementary School"),]
high.educated <- high.educated[!(high.educated$education == "Secondary School"),]
high.educated <- high.educated[!(high.educated$education == "High School"),]

# All participants whose education level is lower than *high school* or equal;
low.educated <- dataset[!(dataset$education == "Associate/Bachelor's Degree"),]
low.educated <- low.educated[!(low.educated$education == "Master Degree or above"),]
```

---

### End Points of Confidence Intervals

At first, I want to show confidence intervals of mean of total.score with *0.9*, *0.95* and *0.99* levels for **high educated** participants.

* **Level of 0.9: **
```{r}
cat(paste0("Lower end point: ", t.test(high.educated$total.score,
                                       conf.level = 0.9)$conf.int[1],
           "\nUpper end point: ", t.test(high.educated$total.score,
                                         conf.level = 0.9)$conf.int[2]))
```

* **Level of 0.95: **
```{r}
cat(paste0("Lower end point: ", t.test(high.educated$total.score,
                                       conf.level = 0.95)$conf.int[1],
           "\nUpper end point: ", t.test(high.educated$total.score,
                                         conf.level = 0.95)$conf.int[2]))
```

* **Level of 0.99: **
```{r}
cat(paste0("Lower end point: ", t.test(high.educated$total.score,
                                       conf.level = 0.99)$conf.int[1],
           "\nUpper end point: ", t.test(high.educated$total.score,
                                         conf.level = 0.99)$conf.int[2]))
```

---

At now, I want to show confidence intervals of mean of total.score with *0.9*, *0.95* and *0.99* levels for **low educated** participants.

* **Level of 0.9: **
```{r}
cat(paste0("Lower end point: ", t.test(low.educated$total.score,
                                       conf.level = 0.9)$conf.int[1],
           "\nUpper end point: ", t.test(low.educated$total.score,
                                         conf.level = 0.9)$conf.int[2]))
```

* **Level of 0.95: **
```{r}
cat(paste0("Lower end point: ", t.test(low.educated$total.score,
                                       conf.level = 0.95)$conf.int[1],
           "\nUpper end point: ", t.test(low.educated$total.score,
                                         conf.level = 0.95)$conf.int[2]))
```

* **Level of 0.99: **
```{r}
cat(paste0("Lower end point: ", t.test(low.educated$total.score,
                                       conf.level = 0.99)$conf.int[1],
           "\nUpper end point: ", t.test(low.educated$total.score,
                                         conf.level = 0.99)$conf.int[2]))
```

---

Finally, I want to show confidence intervals of mean **difference** of total.score with *0.9*, *0.95* and *0.99* levels for **high educated** and **low educated** participants.

* **Level of 0.9: **
```{r}
cat(paste0("Lower end point: ", t.test(high.educated$total.score,
                                       low.educated$total.score,
                                       conf.level = .9)$conf.int[1],
           "\nUpper end point: ", t.test(high.educated$total.score,
                                         low.educated$total.score,
                                         conf.level = .9)$conf.int[2]))
```

* **Level of 0.95: **
```{r}
cat(paste0("Lower end point: ", t.test(high.educated$total.score,
                                       low.educated$total.score,
                                       conf.level = .95)$conf.int[1],
           "\nUpper end point: ", t.test(high.educated$total.score,
                                         low.educated$total.score,
                                         conf.level = .95)$conf.int[2]))
```

* **Level of 0.99: **
```{r}
cat(paste0("Lower end point: ", t.test(high.educated$total.score,
                                       low.educated$total.score,
                                       conf.level = .99)$conf.int[1],
           "\nUpper end point: ", t.test(high.educated$total.score,
                                         low.educated$total.score,
                                         conf.level = .99)$conf.int[2]))
```

---

## Hypothesis Testing

### Pre-processing

I want to remove the rows what have *'I do not want to state'* for gender. Then, I will create a new dataframe. And then, I will regulate factor levels of gender. So,
```{r}
dataset.for.gender <- dataset[!(dataset$gender == "I do not want to state"),]

dataset.for.gender$gender <- factor(dataset.for.gender$gender)
```

I want to remove the rows what have *'Divorced/Widowed'* for marital status. Then, I will create a new dataframe. And then, I will regulate factor levels of marital status. So,
```{r}
dataset.for.maritalstatus <- dataset[!(dataset$marital.status == "Divorced/Widowed"),]

dataset.for.maritalstatus$marital.status <- factor(dataset.for.maritalstatus$marital.status)
```

I want to change level of education with **low** and **high**.

* Low levels;
  + Elementary School
  + Secondary School
  + High School
  
* High levels;
  + Associate/Bachelor's Degree
  + Master Degree or above

So,
```{r}
dataset.for.education <- dataset

dataset.for.education <- dataset.for.education %>%
  mutate(education = recode(education,
                            "Elementary School" = "Low Educated",
                            "Secondary School" = "Low Educated",
                            "High School" = "Low Educated",
                            "Associate/Bachelor's Degree" = "High Educated",
                            "Master Degree or above" = "High Educated"))
```

**Final Steps of Pre-processing**

* Filter for Gender
```{r}
males <- filter(dataset.for.gender, gender == "Male")
females <- filter(dataset.for.gender, gender == "Female")
```

* Filter for Marital Status
```{r}
single <- filter(dataset.for.maritalstatus, marital.status == "Single")
married <- filter(dataset.for.maritalstatus, marital.status == "Married")
```

* Filter for Education
  + **low.educated** and **high.educated** dataframes are created before.


### First Hypothesis
Are there significant difference between mean total.score of **male** and **female**?

**STEPS**

1. **Normality Check;**
```{r}
shapiro.test(males$total.score)
```
**Normality: ** p-value is less than 0.05 and so it is not normal, but I will assume that it is normal.

```{r}
shapiro.test(females$total.score)

```
**Normality: ** p-value is less than 0.05 and so it is not normal, but I will assume that it is normal.

2. **Dependency Check;**

I have two independent classes, so the sampling units(i.e. male and female participants) are totally different. So, I will conduct an independent t-test.

3. **Variance Homogenity**

Ho: sigmaMALE = sigmaFEMALE

Ha: sigmaMALE != sigmaFEMALE
```{r}
var.test(males$total.score, females$total.score)
```
**Homogenity: ** p-value is greater than 0.05 and so null hypothesis for variance homogenity does not reject. In other words, *variances are homogenious.*

4. **Test the Hypothesis**

Ho: muMALE = muFEMALE

Ha: muMALE != muFEMALE
```{r}
t.test(males$total.score, females$total.score,
       paired = FALSE,
       conf.level = .95,
       alternative = "two.sided",
       var.equal = TRUE)
```
**Rejection: ** p-value is greater than 0.05 and the null hypothesis rejection is failed.

**Result: ** The difference between mean total.score of *male* and *female* is not highly significant. In other words, there is not a major difference between them. Even, mean score of male is almost equal to female.


### Second Hypothesis
Are there significant difference between mean total.score of **single** and **married**?

**STEPS**

1. **Normality Check;**
```{r}
shapiro.test(single$total.score)
```
**Normality: ** p-value is less than 0.05 and so it is not normal, but I will assume that it is normal.

```{r}
shapiro.test(married$total.score)

```
**Normality: ** p-value is less than 0.05 and so it is not normal, but I will assume that it is normal.

2. **Dependency Check;**

I have two independent classes, so the sampling units(i.e. single and married participants) are totally different. So, I will conduct an independent t-test.

3. **Variance Homogenity**

Ho: sigmaSINGLE = sigmaMARRIED

Ha: sigmaSINGLE != sigmaMARRIED
```{r}
var.test(single$total.score, married$total.score)
```
**Homogenity: ** p-value is greater than 0.05 and so null hypothesis for variance homogenity does not reject. In other words, *variances are homogenious.*

4. **Test the Hypothesis**

Ho: muSINGLE = muMARRIED

Ha: muSINGLE != muMARRIED
```{r}
t.test(single$total.score, married$total.score,
       paired = FALSE,
       conf.level = .95,
       alternative = "two.sided",
       var.equal = TRUE)
```
**Rejection: ** p-value is greater than 0.05 and the null hypothesis rejection is failed.

**Result: ** The difference between mean total.score of *single* and *married* is not highly significant. In other words, there is not a major difference between them. Even, mean score of single is almost equal to married.


### Third Hypothesis
Are there significant difference between mean total.score of **low educated** and **high educated**?

**STEPS**

1. **Normality Check;**
```{r}
shapiro.test(low.educated$total.score)
```
**Normality: ** p-value is less than 0.05 and so it is not normal, but I will assume that it is normal.

```{r}
shapiro.test(high.educated$total.score)

```
**Normality: ** p-value is less than 0.05 and so it is not normal, but I will assume that it is normal.

2. **Dependency Check;**

I have two independent classes, so the sampling units(i.e. male and female participants) are totally different. So, I will conduct an independent t-test.

3. **Variance Homogenity**

Ho: sigmaLOWEDUCATED = sigmaHIGHEDUCATED

Ha: sigmaLOWEDUCATED != sigmaHIGHEDUCATED
```{r}
var.test(low.educated$total.score, high.educated$total.score)
```
**Homogenity: ** p-value is greater than 0.05 and so null hypothesis for variance homogenity does not reject. In other words, *variances are homogenious.*

4. **Test the Hypothesis**

Ho: muMALE = muFEMALE

Ha: muMALE != muFEMALE
```{r}
t.test(low.educated$total.score, high.educated$total.score,
       paired = FALSE,
       conf.level = .95,
       alternative = "two.sided",
       var.equal = TRUE)
```
**Rejection: ** p-value is less than 0.05 and the null hypothesis rejection is not failed.

**Result: ** The difference between mean total.score of *male* and *female* is highly significant. In other words, there is a major difference between them.

---

## ANOVA or Regression

To analyse hypothesis rejection, I've used t-test function for two levels. Now, I will analyse for three or more levels and I cannot use t-test again. So, I will use ANOVA.

### Are there significant differences between mean of total.score of different education levels?

**STEPS**

1. **Visualization**

First of all, I want to show total.score informations based on levels of *education*.
```{r}
boxplot(data = dataset, total.score ~ education)
```

As you can see easily, there are different education levels and their score results are different, too.

2. **Normality Check;**

At now, I will check the normality for every levels of education.
```{r}
shapiro.test(filter(dataset, education == "Elementary School")$total.score)
```
**Normality: ** p-value is greater than 0.05 and so it is normal.

```{r}
shapiro.test(filter(dataset, education == "Secondary School")$total.score)
```
**Normality: ** p-value is greater than 0.05 and so it is normal.

```{r}
shapiro.test(filter(dataset, education == "High School")$total.score)
```
**Normality: ** p-value is less than 0.05 and so it is not normal, but I will assume that it is normal.

```{r}
shapiro.test(filter(dataset, education == "Associate/Bachelor's Degree")$total.score)
```
**Normality: ** p-value is less than 0.05 and so it is not normal, but I will assume that it is normal.

```{r}
shapiro.test(filter(dataset, education == "Master Degree or above")$total.score)
```
**Normality: ** p-value is less than 0.05 and so it is not normal, but I will assume that it is normal.

3. **Variance Homogenity**

Afterwards, I will check the variance homogenity and I will create a null hypothesis which is assumed equality of variance of every education levels.

Ho: varELEMENTARY = varSECONDARY = varHIGH = varASSOCIATE/BACHELOR = varMASTER

Ha: At least one of them is different
```{r}
bartlett.test(data = dataset, total.score ~ education)
```
**Homogenity: ** p-value is greater than 0.05 and so null hypothesis for variance homogenity does not reject. In other words, *variances are homogenious.*

4. **Analysis of Variance**

After homogenity checking process, I will analyse the variances.
```{r}
model.education <- aov(data = dataset, total.score ~ education)
model.education$coefficients
```

Intercept is mean total.score of participants who are stated their education level is *Elementary School* and its value is **80**. Then, the mean total.score of other levels;

Education Level             | Mean total.score
---------------             | ----------------
Secondary School            | 80 - 4.545455 = 75.45454
High School                 | 80 - 4.25 = 75.75
Associate/Bachelor's Degree | 80 + 1.928571 = 81.92857
Master Degree or above      | 80 - 12.5 = 67.5


Now, I will show a summary of model for education levels. Also, create a null hypothesis and alternative hypothesis:

Ho: muELEMENTARY = muSECONDARY = muHIGH = muASSOCIATE/BACHELOR = muMASTER

Ha: At least one of them is different
```{r}
summary(model.education)
```

From output results, I will use F-value and p-value.

* F-value
  + It means effect of education levels is 3.096 times important.
  + I will find the expected value using qf() function and analyse them for null hypothesis rejection.

* p-value
  + It is for rejection of null hypothesis.
  + Null hypothesis rejects if p-value is less than 0.05.
  + On the other hand, null hypothesis will be failed to reject.
  
  In this output result, p-value is 0.0171 and it is less than 0.05. So, the null hypothesis rejects. In other words, at least one level's mean total.score is different from others.

Now, I utilize qf() function and it gives expected value of F-value.
```{r}
qf(1-0.05, 4, 195)
```

The expected F-value is approximate 2.418 and it is less than F-value of education which is 3.086. So, null hypothesis should be rejected. I knew from p-value processes, null hypothesis is rejected and this result proves it.


5. **Find different one(s) out**
```{r}
TukeyHSD(model.education)
```

